---
title: "Lavaan"
author: "Sith"
date: "2023-03-30"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
setwd('C:/Users/lfult/Documents/Minimum Wage/')

```


# Lavaan

Minimum set of libraries

```{r}

require(car)
require(ggcorrplot)
require(kableExtra)
require(lavaan)
require(lavaanPlot)
require(psych)
require(ResourceSelection)
require(semPlot)
require(tidySEM)
require(performance)

mydata=read.csv('jasp4.csv')

```

# Functions

Correlation, printing, and citing

```{r}


corfunction=function(d,m='pearson',sz=2){
  mycorr=cor(d[, 1:ncol(d)], method=m); p.mat=ggcorrplot::cor_pmat(d[,1:ncol(d)])
  myplot=ggcorrplot(mycorr, hc.order=T,type="lower",
                    colors=c("red", "white","green"),tl.cex = 8, 
                    tl.col = "black", lab=TRUE, lab_size=sz, p.mat=p.mat,sig.level=0.01,
                    insig="pch", pch=4)
  print(myplot)}

myprint=function(x){x%>%kbl()%>%kable_classic(html_font = "Cambria")}
mycite=function(x){citation(x)}

```



# Descriptives

Descriptives prior to normalizing


```{r}

myprint(describe(mydata))


```

# Correlations

```{r}
#colnames(mydata)[4:13]=c('X1A','X1B','X1C','X2A','X2B','X2C','X3A', 'X3B','X3C', 'New3A')
corfunction(mydata[,-c(1:3)], ,3)

```


# KDE Pairs Plot 1

KDE Pairs of data

```{r}

kdepairs(mydata[, 2:3])
kdepairs(mydata[,4:6])
kdepairs(mydata[,7:9])
kdepairs(mydata[,10:13])


```


# Normalize

Z-scores to adjust for magnitudes


```{r}

tempname=mydata$GeoName
mydata[,4:13]=as.data.frame(apply(mydata[, 4:13], 2, 'scale'))
mydata$GeoName=tempname
myprint(describe(mydata))

```



# Correlations

Not hierarchically clustered


```{r}

corfunction(mydata[,4:13])

```


# M1 Base with Cov

Baseline model

Estimate of negative variance exists for the forced additive measure of Economic Freedom (P2 and P3). 


```{r}

colnames(mydata)[4:13]=c('X1A','X1B','X1C','X2A','X2B','X2C','X3A', 'X3B','X3C', 'New3A')

mod='
#Score~EF
Economic_Freedom=~1/3*Government_Spending+1/3*Taxation+1/3*Labor_Market_Freedom
Government_Spending=~1/3*X1A+1/3*X1B+1/3*X1C
Taxation=~1/3*X2A+1/3*X2B+1/3*X2C
Labor_Market_Freedom=~1/3*X3A+1/3*X3B+1/3*X3C

'


fit=cfa(mod, data=mydata, do.fit=T, estimator='DWLS')
(mys1=summary(fit, standardized=T,fit.measures = TRUE))
lavaanPlot(model = fit, edge_options = list(color = "grey"),coefs = TRUE, covs=F)


temp2=lavPredict(fit)[,1]
mydata$ScoreM1=10*(temp2-min(temp2))/(max(temp2)-min(temp2)) #scale between 0 and 10


```

# M2 Sub with Cov

Estimate of negative variance exists for the forced additive measure of Economic Freedom (P2 and P3). 

```{r}


mod='
Economic_Freedom=~1/3*Government_Spending+1/3*Taxation+1/3*Labor_Market_Freedom
Government_Spending=~1/3*X1A+1/3*X1B+1/3*X1C
Taxation=~1/3*X2A+1/3*X2B+1/3*X2C
Labor_Market_Freedom=~1/3*New3A+1/3*X3B+1/3*X3C

'

fit=cfa(mod, data=mydata, do.fit=T, estimator='DWLS')
(mys2=summary(fit, standardized=T,fit.measures = TRUE))
lavaanPlot(model = fit, edge_options = list(color = "grey"),coefs = TRUE,covs=T)



temp2=lavPredict(fit)[,1]
mydata$ScoreM2=10*(temp2-min(temp2))/(max(temp2)-min(temp2)) #scale between 0 and 10

```

# M3 Non-Fixed Weights

Statistically significant estimate of negative variance exists for the forced additive measure of Economic Freedom (P1). 

```{r}

lavOptions("verbose")

mod='
Economic_Freedom=~Government_Spending+Taxation+Labor_Market_Freedom
Government_Spending=~X1A+X1B+X1C
Taxation=~X2A+X2B+X2C
Labor_Market_Freedom=~X3A+X3B+X3C


'

fit=cfa(mod, data=mydata, do.fit=T, estimator='DWLS')
(mys3=summary(fit, standardized=T, fit.measures = TRUE))
lavaanPlot(model = fit, edge_options = list(color = "grey"),coefs = TRUE, covs=F)



temp2=lavPredict(fit)[,1]
mydata$ScoreM3=10*(temp2-min(temp2))/(max(temp2)-min(temp2)) #scale between 0 and 10

```


# M4 Sub Weights Free

Statistically significant estimate of negative variance exists for the forced additive measure of Economic Freedom (P2). 

```{r}


mod='
Economic_Freedom=~Government_Spending+Taxation+Labor_Market_Freedom
Government_Spending=~X1A+X1B+X1C
Taxation=~X2A+X2B+X2C
Labor_Market_Freedom=~New3A+X3B+X3C

'

fit=cfa(mod, data=mydata, do.fit=T, estimator='DWLS')
(mys4=summary(fit, standardized=T, fit.measures=TRUE))
lavaanPlot(model = fit, edge_options = list(color = "grey"),coefs = TRUE,covs=F)



temp2=lavPredict(fit)[,1]
mydata$ScoreM4=10*(temp2-min(temp2))/(max(temp2)-min(temp2)) #scale between 0 and 10

```

# EFA 1

```{r}

mod='
Economic_Freedom=~X1A+X1B+X1C+X2A+X2B+X2C+X3A+X3B+X3C
'

fit=efa(data=mydata[,4:12], do.fit=T, estimator='DWLS', rotation='varimax', nfactors=3, sample.cov=cor(mydata[,4:12]), rotation.args = list(geomin.epsilon = 0.05, rstarts = 1, orthogonal=T))
summary(fit, standardized=T, fit.measures = TRUE, cutoff=0.4)



```



# M5 Base, EFA-Based with Covariance Structure

Model fits.  (Estimate of negative variance statistically insignificant) 

```{r}

mod='
#Score~EF
Economic_Freedom=~F1+F2+F3
F1=~X2A+X2B
F2=~X1A+X1B+X3A+X3B
F3=~X1C+X3C


'
#optimal covariance structure

fit=sem(mod, data=mydata, do.fit=T, estimator='DWLS')
(mys5=summary(fit, standardized=T, fit.measures=T))
lavaanPlot(model = fit, edge_options = list(color = "grey"),coefs = TRUE, covs=F)

temp2=lavPredict(fit)[,1]
mydata$ScoreM5=10*(temp2-min(temp2))/(max(temp2)-min(temp2)) #scale between 0 and 10

```
# EFA 2

```{r}

mod='
Economic_Freedom=~X1A+X1B+X1C+X2A+X2B+X2C+New3A+X3B+X3C

'

fit=efa(data=mydata[, c(4:9,11:13)], do.fit=T, estimator='DWLS', rotation='varimax', nfactors=3, sample.cov=cor(mydata[, c(4:9,11:13)]), rotation.args = list(geomin.epsilon = 0.05, rstarts = 1, orthogonal=T))
summary(fit, standardized=T, fit.measures = TRUE, cutoff=.4)



```



# M6 Sub, EFA-Based, with Covariance Structure

Model fits.  (Estimate of negative variance statistically insignificant) 


```{r}


mod='
Economic_Freedom=~F1+F2+F3
F1=~X1A+X1B+X3B
F2=~X2A+X2B
F3=~X1C+X3C+New3A


'
#optimal covariance structure for fit

fit=sem(mod, data=mydata, do.fit=T, estimator='DWLS')
(mys6=summary(fit, standardized=T, fit.measures=T))
lavaanPlot(model = fit, edge_options = list(color = "grey"),coefs = TRUE, covs=F)



temp2=lavPredict(fit)[,1]
mydata$ScoreM6=10*(temp2-min(temp2))/(max(temp2)-min(temp2)) #scale between 0 and 10

```


# Top 10

```{r}

newdata=subset(mydata, select=c('GeoName','MEFI_Score', 'New_MEFI_Score', 'ScoreM1', 'ScoreM2', 'ScoreM3','ScoreM4','ScoreM5','ScoreM6'))

MEFI=newdata[order(-newdata$MEFI_Score),]
MEFI$RankOriginal=seq(1:nrow(MEFI))

MEFI=MEFI[order(-MEFI$New_MEFI_Score),]
MEFI$RankNewScore=seq(1:nrow(MEFI))

MEFI=MEFI[order(-MEFI$ScoreM1),]
MEFI$RankM1=seq(1:nrow(MEFI))

MEFI=MEFI[order(-MEFI$ScoreM2),]
MEFI$RankM2=seq(1:nrow(MEFI))

MEFI=MEFI[order(-MEFI$ScoreM3),]
MEFI$RankM3=seq(1:nrow(MEFI))

MEFI=MEFI[order(-MEFI$ScoreM4),]
MEFI$RankM4=seq(1:nrow(MEFI))

MEFI=MEFI[order(-MEFI$ScoreM5),]
MEFI$RankM5=seq(1:nrow(MEFI))

MEFI=MEFI[order(-MEFI$ScoreM6),]
MEFI$RankM6=seq(1:nrow(MEFI))

write.csv(MEFI,'MEFI.csv', row.names=F)


```


# Rank Correlation

```{r}

cor(MEFI[, c(10:11,14:17)], method='spearman')
corfunction(MEFI[,c(10:11,14:17)], 'spearman', 3)

```

# Model Performance

```{r}

require(tidyverse)
mydf=as.data.frame(round(rbind(mys1$fit, mys2$fit, mys3$fit,mys4$fit,mys5$fit,mys6$fit),5))
rownames(mydf)=c('M1','M2', 'M3','M4','M5','M6')
mydf$Model=rownames(mydf)
mydf[,6:8]%>%kbl(caption='Scaled Chi Squared Test of Fit (Null:  Perfect Fit)')%>%kable_classic(html_font='Cambria')
mydf[,c(11:13)]%>%kbl(caption='Tests of RMSEA')%>%kable_classic(html_font='Cambria')

mysub=mydf[,c(9,10)]%>%pivot_longer(everything(), names_to='metric', values_to='value')
mysub$Model=c(rep('M1',2),rep('M2',2),rep('M3',2),rep('M4',2),rep('M5',2),rep('M6',2))


myplot=ggplot(mysub, aes(x=metric, y=value, fill=Model))+geom_bar(stat='identity', alpha=.5, position='dodge')+
  geom_text(aes(label = round(value, 3)), 
            position = position_dodge(.9),
            color="black",vjust = .5,hjust =1, angle = 90)+xlab("")+ylab("Metric Value")
myplot

path1 <- data.frame(x=c(.5,1.5),y=c(.95,.95), Model=c('Benchmark'))
path2 <- data.frame(x=c(1.5,2.5),y=c(.9,.9), Model=c('Benchmark'))
q <- myplot + geom_path(data=path1,aes(x=x, y=y, col=Model))
q <- q + geom_path(data=path2,aes(x=x, y=y, col=Model))+ylim(0, 1.2)
q



```



# Top k


```{r}
k=10

BaseTop=MEFI[MEFI$RankOriginal<=k,]
BaseTop=select(BaseTop, c('GeoName','RankOriginal'))
BaseTop=BaseTop[order(BaseTop$RankOriginal),"GeoName"]

SubTop=MEFI[MEFI$RankNewScore<=k,]
SubTop=select(SubTop, c('GeoName','RankNewScore'))
SubTop=SubTop[order(SubTop$RankNewScore),"GeoName"]


M3Top=MEFI[MEFI$RankM3<=k,]
M3Top=select(M3Top, c('GeoName','RankM3'))
M3Top=M3Top[order(M3Top$RankM3),"GeoName"]

M4Top=MEFI[MEFI$RankM4<=k,]
M4Top=select(M4Top, c('GeoName','RankM4'))
M4Top=M4Top[order(M4Top$RankM4),"GeoName"]

M5Top=MEFI[MEFI$RankM5<=k,]
M5Top=select(M5Top, c('GeoName','RankM5'))
M5Top=M5Top[order(M5Top$RankM5),"GeoName"]

M6Top=MEFI[MEFI$RankM6<=k,]
M6Top=select(M6Top, c('GeoName','RankM6'))
M6Top=M6Top[order(M6Top$RankM6),"GeoName"]

newdf=as.data.frame(cbind(BaseTop, SubTop, M3Top, M4Top, M5Top,M6Top))
myprint(newdf)
myt=table(as.matrix(newdf))
myprint(myt[order(myt, decreasing=T)])

```

# Bottom k


```{r}

k=nrow(MEFI)-10

BaseBottom=MEFI[MEFI$RankOriginal>k,]
BaseBottom=select(BaseBottom, c('GeoName','RankOriginal'))
BaseBottom=BaseBottom[order(BaseBottom$RankOriginal),"GeoName"]

SubBottom=MEFI[MEFI$RankNewScore>k,]
SubBottom=select(SubBottom, c('GeoName','RankNewScore'))
SubBottom=SubBottom[order(SubBottom$RankNewScore),"GeoName"]


M3Bottom=MEFI[MEFI$RankM3>k,]
M3Bottom=select(M3Bottom, c('GeoName','RankM3'))
M3Bottom=M3Bottom[order(M3Bottom$RankM3),"GeoName"]

M4Bottom=MEFI[MEFI$RankM4>k,]
M4Bottom=select(M4Bottom, c('GeoName','RankM4'))
M4Bottom=M4Bottom[order(M4Bottom$RankM4),"GeoName"]

M5Bottom=MEFI[MEFI$RankM5>k,]
M5Bottom=select(M5Bottom, c('GeoName','RankM5'))
M5Bottom=M5Bottom[order(M5Bottom$RankM5),"GeoName"]

M6Bottom=MEFI[MEFI$RankM6>k,]
M6Bottom=select(M6Bottom, c('GeoName','RankM6'))
M6Bottom=M6Bottom[order(M6Bottom$RankM6),"GeoName"]

newdf=as.data.frame(cbind(BaseBottom, SubBottom, M3Bottom, M4Bottom, M5Bottom,M6Bottom))
myprint(newdf)
myt=table(as.matrix(newdf))
myprint(myt[order(myt, decreasing=T)])


```

# Best Fit Model

```{r}

myprint(MEFI[order(MEFI$RankM6, decreasing=F),c(1,17)])

```



